---
title: "Quantified Self"
output:
  html_document: default
  html_notebook: default
---

## Setup:

```{r}

```
---

## Loading and preprocessing the data

```{r}
steps <- read.csv('activity.csv')
steps$date <- as.Date(steps$date)
```


```{r}
str(steps)
```


```{r}
summary(steps)
```


---

## What is mean total number of steps taken per day?
Ignoring the missing values in the dataset.

```{r}
data <- aggregate(steps~date, data=steps, sum, na.rm=T)
hist(data$steps, main='Total Steps', xlab='steps')
```


```{r eval=FALSE}
mean(data$steps)
median(data$steps)
```

Mean steps per day is **`r mean(data$steps)`**

Median steps per day is **`r median(data$steps)`**

---

## What is the average daily activity pattern?

```{r}
data <- aggregate(steps~interval, data=steps, mean, na.rm=T)
max_steps <- max(data$steps)
interval <- data[data$steps==max_steps, 1 ]
with( data, 
  plot( interval, steps, 
        type='l',
        ylab='avg steps'
  )  )
abline(v=interval, col='red')
```

The 5-minute interval num **`r interval`**, on average across all the days in the dataset, contains the maximum number of steps: **`r max_steps`**

---

## Imputing missing values

```{r eval=FALSE}
sum(is.na(steps$steps))
```

The total number of missing values in the dataset is **`r sum(is.na(steps$steps))`**.

My strategy for filling in all of the missing values in the dataset:

1. Start by computing the means for each interval with NAs removed.
2. Replace any NA in the data set with the computed mean.
```{r}
means <- aggregate(steps ~ interval, data=steps, mean, na.rm=T)

steps2 <- steps
steps2$steps <- ifelse(
            is.na(steps2$steps)
            , means[ match(steps2$interval, means$interval), 2]
            , steps2$steps  )
```

After imputation the updated histogram is:
```{r}
data <- aggregate(steps~date, data=steps, sum, na.rm=T)
hist(data$steps, main='Total Steps', xlab='steps')
```

The new values for median and mean are:
```{r eval=FALSE}
mean(data$steps)
median(data$steps)
```

Mean steps per day is **`r mean(data$steps)`**

Median steps per day is **`r median(data$steps)`**

The impact of imputing missing data seems **negligable**.

---

## Are there differences in activity patterns between weekdays and weekends?

```{r}
steps2$weekday <- as.factor(weekdays(steps2$date))
steps2$dayType <- ifelse(steps2$weekday %in% c('Saturday','Sunday'), 'weekend', 'weekday')
unique(steps2[,c(4,5)])
```

```{r}
par(mfrow=c(2,1))
par(mar=c(2,1,1,1))
data <- subset(steps2, dayType=='weekend')
with(data, plot(interval, steps, type='l', main='weekend') )

data <- subset(steps2, dayType=='weekday')
with(data, plot(interval, steps, type='l', main='weekday') )
```

Yes, there are differences in activity patterns between weekdays and weekends. The activity pattern for weekdays reflects an earlier start to the day than on the weekend.